Stochastic Variance Reduced Policy Gradient

### 2018-01-25 svrpg.tex
added minibatch to proofs and theorem

\begin{assumption}\label{ass:M2}
	There is a constant $\Omega<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq \Omega
\]
\end{assumption}

@@

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $F>0$ and for proper choice of step sizes $\alpha_t$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} \\
		&\qquad+
		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
		+\frac{F\Omega}{\psi}
\end{align*}
\end{restatable}

@@

+dG^2R^2\left(M_2-1\right)

@@

			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -

@@

			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)

@@

	\mathbb{E}_{t\mid s}\left[\left\|\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)

@@

		- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right] 

@@

			-\Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}\right\|^2\right]  \nonumber\\

@@

	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\

@@

	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)

@@

	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 

@@

	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 

@@

@@

	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\Ets{(\omega(\tau_i)-1)^2} \\

@@

	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\

@@

	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2\Omega}{(1-\gamma)^2},

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
	+\frac{f_t\Omega}{\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_tL^2+\frac{\alpha_t^2L^3}{2}, \\
	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
\end{align*}
In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
\begin{align*}
&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
&\beta_t > c_{t+1}.
\end{align*}
\end{restatable}
\begin{proof}
	We have:
	\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\geq
	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
\end{align}
where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.

@@

	&\left.\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\Omega\right)

@@

	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_tL^2+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\

@@

	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2\Omega}{2(1-\gamma)^2} \nonumber\\

@@

	-f_t\Omega,\nonumber

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\Psi_t, \\
& D \coloneqq \max_t d_t \\
& F \coloneqq \max_t f_t.
\end{align*}
The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.

>>>>

\begin{assumption}\label{ass:M2}
	There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq W
\]
\end{assumption}

\begin{assumption}\label{ass:REINFORCE}
There is a constant $V<\infty$ such that, for each policy $\pol$:
\[
	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq V
\]
\end{assumption}

@@

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,D, F$ and for proper choice of step sizes $\alpha_t$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{Z}{N}
		+\frac{\Omega}{b}
\end{align*}
\end{restatable}

@@

+dG^2R^2W

@@

			+\frac{1}{b}\sum_{i=0}^{b-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 
			-\frac{1}{b}\sum_{i=0}^{b-1}

@@

			+\frac{1}{b}\sum_{i=0}^{b-1}\left( 
			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)

@@

	\mathbb{E}_{t\mid s}\left[\left\|
		\frac{1}{b}\sum_{i=0}^{b-1}\left(
		\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)

@@

		- \Ets{
			\frac{1}{b}\sum_{i=0}^{b-1}\left(
			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 

@@

		- \Ets{
			\frac{1}{b}\sum_{i=0}^{b-1}\left(
			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 
		\nonumber\\%
	&\qquad+\Ets{\norm[]{
			\frac{1}{b}\sum_{i=0}^{b-1}\left(
			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \\

@@

			\frac{1}{b^2}\sum_{i=0}^{b-1}
			\Ets{\norm[]{
			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -

@@

			\frac{1}{b^2}\sum_{i=0}^{b-1}
			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)

@@

			+\frac{1}{b^2}\sum_{i=0}^{b-1}
			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 

@@

	+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}\nonumber\\
		\frac{1}{b^2}\sum_{i=0}^{b-1}
		\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 

@@

% ...
% ...

@@

	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{b^2}\sum_{i=0}^{b-1}\Ets{(\omega(\tau_i)-1)^2} \\

@@

	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{b-1}\Varts{\omega(\tau_i)} \nonumber\\

@@

	&\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2W}{(1-\gamma)^2b},

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_t\Tr\Varts{\gradApp{\tilde{\vtheta}^s}{1}}}{N\Psi_t}
	+\frac{f_tW}{b\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}, \\
	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
\end{align*}
In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
\begin{align*}
&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
&\beta_t > c_{t+1}.
\end{align*}
\end{restatable}
\begin{proof}
	We have:
	\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\geq
	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2},
\end{align}
where (11) is from the L-smoothness of $J(\vtheta)$, (12) is from the SVRPG update, and (13) is from Lemma \ref{lemma:aux0}.

@@

	&\left.\qquad+\frac{L^2}{b}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2b}W\right)

@@

	& = \Ets{J(\vtheta_t^{s+1}) - \left(c_{t+1}\left(\frac{3}{2}+\alpha_t\beta+\alpha_t^2\right)+\alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}\right)\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}}^2} \nonumber\\

@@

	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})G^2R^2W}{2(1-\gamma)^2b} \nonumber\\

@@

	-\frac{f_t}{b}W,\nonumber

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\{\Psi_t\}, \\
& Z \coloneqq \frac{\max_t\{d_t\}V}{\psi} \\
& \Omega \coloneqq \frac{\max_t\{f_t\}W}{\psi}.
\end{align*}
The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
